A Low-cost Natural Interaction System based on a Camera Hand-Gestures Recognizer

### 2011-02-04 paper.tex
Last Franck and Thomas notes

\frontmatter          % for the preliminaries
\pagestyle{headings}  % switches on printing of running heads
\mainmatter              % start of the contributions
%
\title{A Low-cost Natural Interaction System based on a Camera Hand-Gestures Recognizer}
\author{Mohamed-Ikbel Boulabiar\inst{1} \and Thomas Burger\inst{2}
Franck Poirier\inst{3} \and Gilles Coppin\inst{1}}
%
\institute{LAB-STICC, Telecom-Bretagne, France,\\
\email{boulabiar@gmail.com, gilles.coppin@telecom-bretagne.eu},
\and
LAB-STICC, University of Bretagne-Sud, France
\email{thomas.burger@univ-ubs.fr}
\and
VALORIA, University of Bretagne-Sud, France
\email{franck.poirier@univ-ubs.fr}
}

@@

\section{Introduction}
HCI research focuses more attention than before on enhancing the user
experience regarding human-display interaction.
This area of research and specifically gesture, work on making interact natural
without the need of clicking on buttons and touching screens with first systems
since the 1980's \cite{Bolt1980}, \cite{Pavlovic97visinter}, \cite{Derpanisreview}.
However, in everyday life, interaction with home devices having
embedded computational power, such as interactive TVs,
one does not make benefit from these new interaction paradigms yet.

We present in this paper an approach that could allow the detection and the
interpretation of gestures of a TV spectator from a simple and low cost camera. 
RevTV is the name of the French project for adding new interaction techniques
to TVs to allow the inclusion of the telespectator.

The final objective of this study is to control the animations of an avatar
that should be inserted within a TV show beside the presenter or actor.
One major scenario that we rely on, corresponds to some educational game where a pupil
controls his/her avatar which interact in a forecast program led by a ``real'' TV animator.
During such kind of scenarios, commands interaction
like pointing, selecting and moving, as well as natural gesture animation, are required.

% Il faut rajouter une description du projet Rev-TV
%Le dernier paragraphe n'est pas une description ?

\section{Gesture Semantics}
Symbolic gestures, such as emblems, play an important role in human
communication as they can fully take the place of words.
These gestures and spoken language are processed by the same system in the human brain.

Gestures do not come alone, they are combined with speech, or with cognitive
activities which can be detected from the interaction between two human in a
discussion, these type of gestures are used to add more precision on the details
and they are made by the same areas of the cortex \cite{SymbolicGest}.

@@

We try to extract some semantics of gestures by the analysis of the movements of
a user explaining a story and searching for a relation between gestures and the
part of the story being told \cite{gestureThought}.
%The result of this is the extraction of gesticulations .

After extraction, most of gestures semantics can be divided into two parts for
their possible future use as:
\begin{description}
 \item[Animation Gestures] used to animate a virtual avatar.
 \item[Command Gestures] used to launch a specific predefined action.
\end{description}
%%%
% toute cette partie me semble incompréhensible... (entre la debut de la section 2 et la début de la 2.1)
% Je veux bien t'aider à la rédiger, mais je ne comprends pas vraiment ce que tu veux dire... 
% Quel est ton message ?

@@

\subsection{Gestures taxonomies}
A rapid look the possible gestures human can perform \cite{Gesturecraft},
lead to the conclusion that the amount of possibilities is tremendous.
Nonetheless, the gestures can be classified according to various criteria not
only to simplify their recognition, but also to allow their reproducing on an
avatar to putting tags on them for future analysis\cite{gestureRecreation}.
Here, we consider a taxonomy based on the context of the gestures as well as on their types:
%%%%%%%
%By looking at the possible gestures human can generate \cite{Gesturecraft}, we
%get a huge amount of possibilities which depend on the context and on the type.
%So, there is a need for a taxonomy.
%Many possible taxonomies exist according to the technical or analytical aspects:
%In order to describe and analyse each gesture, we can rely on the following aspects:
\ignore{\begin{description}
 \item[Gesture Creation Space:] We use this criteria to distinguish between gestures made in the 2D or 3D space.
 \item[Number of used strokes:] We use this criteria to distinguish between gestures constructed by only one path and between those using many paths, so the recognition needs to wait for all strokes before the recognition
 \item[Gesture context:] We use this criteria to distinguish gesture based on the specific use of it, gestures for disabled people, gestures in talk-shows, gestures in meetings.
\end{description}}
%Gesture and Thoughts, page 18
According to McNail \cite{gestureThought}, it is possible and useful to
distinguish between these gestures and a possible taxonomy can provide us these
types:
\begin{description}
 \item[Gesticulation] is a motion that embodies a meaning reliable to the
accompanying speech. It is made chiefly with the arms and hands but is not
restricted to these body parts.
 \item[Speech-linked gestures] are parts of sentences themselves, the gesture
completes the sentence structure.
 \item[Emblems] are conventionalized signs, such as thumbs-up or the ring (first
finger and thumb tips touching, other fingers extended) for ``OK.''
 \item[Pantomime] is dumb show, a gesture or sequence of gestures conveying a
narrative line, with a story to tell, produced without speech.
\end{description}

In our work, we focalize on emblems for commands and gesticulations, as both of
them are of prime importance for the scenario under consideration.
%We have chosen to analyze gesticulations as they are the most frequent.
%, and project them on the technical taxonomies.
%And we have also made our own extraction from Talk-Shows.
%Nous avons choisi celle-ci, car elle correspond aux besoins que nous avons
%identifié dans les scenarios d'utilisation de la télé interactive

@@

% Il faut rédiger complétement cette partie...
%C'est pas compréhensible, et il n'y a pas de verbe dans la seconde phrase.
In order to identify relevant gestures we have analyzed some talk-shows in order
to extract most frequent gestures being performed.
According to our scenario of use in the context of Interactive TV, we need
gesture to trigger commands but also for natural interaction with the avatar.
We have created a vector based gestures icons to represent them.
These 4 icons in the figure\ref{handg}, which are part of the ``ishara'' library,
represent the software supported gestures in our application.
%and defining a benchmark scenario.
%We have tried to target the following kinds of gestures in a first step.

\png{handg}{Family of supported gestures represented using the free gestures icons library from https://github.com/boulabiar/ishara }

Figure\ref{handg} shows our supported gestures. We provide a movement mode
used to select the 4 directions, then a multitouch, then a command mode, then a hand fingers recognition,
and, finally, a cursor movement mode.
%\pagebreak 
We have selected 4 basic gestures modes to be supported in our system as a first step.
Then to be extended depending on the context fulfilling the requirements
of our scenario.

@@

In the figure \ref{pipeline}, we show the pipeline of the modules used to
extract gestural informations. Each part is described in details in the next
subsections.

@@

\subsection{Used Materials}
We have chosen a low-cost camera for our work that can be integrated with other components in the future.
The PS3Eye camera matched our expectations with a 640x480 resolution combined with a 60 frames-per-second capability.
This camera costs about 40\$, which is a reasonable price for a potential market deployment.

\subsection{Color Transformations} %%%%%%%%%%%%%%%%%%%
The video processing is based on several modules, such as illustrated in Fig ... .
First, a basic but adaptive skin color segmentation of a large interval of human skin varieties is performed \cite{skinColorSeg}.
We also use a histogram equalisation technique to better enlarge the colors to all the space available.
Then we transform the color space from BGR, which come from the camera to YCrCb,
as it is the most adapted one to skin segmentation
Moreover, the corresponding transformation is linear \cite{skinColorSeg}, which helps computation time saving.

Other color spaces like HSV or CIE-Lab can be used for skin color segmentation,
but the transformation from BGR either takes more time or needs more parameters
in the next step.

@@

\section{Blobs Tracking}
After having the hands and head blobs identified, an efficient
tracking method based on the appearance model for occlusion handling\cite{app06}.
We have used the cvblob implementation \cite{cvblob} for that algorithm which 
stores the measures of bounding boxes distance and provide tracks instead of just blobs.  %distantBlobTrack.

When the matrix of distance is set up, the coming loops
are executed to handle blobs changes:
\begin{itemize}
 \item A loop to detect inactive tracks: those tracks with no blobs near.
 \item Detect and create new tracks: those blobs without a track near.
 \item A loop to assign blobs to tracks. It makes clusters with blobs that are close and assign them to a track.
In this step some tracks could merge in one.
 \item A last loop which check all inactive tracks to delete the old ones.
\end{itemize}

To better handle blobs, the bounding path of the blobs
is simplified to a set of few points when needed.
This is used to identify picks for the counter mode.

@@

A phase of calibration is triggered when the logic of tracking is lost or damaged.
We have used four modes in a first core with the possibility to add more sub-gestures in the future.
This core will be used to control a children game having to select, choose, and answer basic questions.
We have stick to that scenario to produce these gestures.

@@

\begin{figure}[!htb]
\centering
  \includegraphics[width=0.45\textwidth]{./pics/point.png}
  \includegraphics[width=0.4\textwidth]{./pics/simplify.png}
  \includegraphics[width=0.46\textwidth]{./pics/fingers.png}
  \includegraphics[width=0.46\textwidth]{./pics/move.png}
  \caption{A sub-list of recognition modes showing pointing, number of hand
fingers counting and the movement mode}
%  \label{fig:Gestes}
\end{figure}

@@

Spacial gestural interactions always lack from the feedback for the interaction.
This affects also the usability because the user can no more be guided in space
as he was with surface movement or simple physical joysticks.

\section{System integration of this interaction and future work}
Human brain perisylvian areas, identified as the core of the brain's language
system, are not in fact committed to language processing, but may function as a
modality-independent semiotic system that plays a broader role in human
communication, linking meaning with symbols whether these are words, gestures,
images, sounds, or objects \cite{SymbolicGest}.
This natural handling of multimodal communications gives an argument to build
recognizer taking part of most of these modalities specially to remove ambiguous
gesture meaning decision cases.

@@

\section*{Conclusion}
In this paper we have shown a proof-of-concept mechanism to recognize hand
gestures then use them to control a TV or to take actions in a Game scenario.
We have also freely provided a library which can be used to represent gestures
in other scenarios.
The evolution of our system is being discussed to add more modalities even if we
are aware of some of the limits \cite{Oviatt99tenmyths}.

>>>>

\frontmatter % for the preliminaries
\pagestyle{empty} % switches on printing of running heads
\mainmatter % start of the contributions
%
\title{A Low-cost Natural Interaction System based on a Camera Hand-Gestures Recognizer}
\author{Mohamed-Ikbel Boulabiar\inst{1} \and Thomas Burger\inst{2}
Franck Poirier\inst{3} \and Gilles Coppin\inst{1}}
%
\institute{LAB-STICC, Telecom-Bretagne, France,\\
\email{boulabiar@gmail.com, gilles.coppin@telecom-bretagne.eu},
\and
LAB-STICC, University of Bretagne-Sud, France
\email{thomas.burger@univ-ubs.fr}
\and
VALORIA, University of Bretagne-Sud, France
\email{franck.poirier@univ-ubs.fr}
}

@@

\section{Introduction}
HCI research focuses more attention than before on enhancing the user
experience regarding human-display interaction.
This area of research focus on making interaction more natural, so that it is not necessary to click on buttons and to touch screens. To do so, automatic gesture recognition is an interesting paradigm \cite{Bolt1980}, \cite{Pavlovic97visinter}, \cite{Derpanisreview}.
However, in everyday life, for interaction with home devices having
embedded computational power, such as interactive TVs, one does not make benefit from these new interaction paradigms yet.

We present in this paper an approach that allows the detection and the
interpretation of gestures of a TV spectator from a simple and low cost camera. this work takes place in the context of the French FUI\footnote{Fond Unique Interministeriel} Project RevTV,  the aim of which is to add new interaction techniques, sos that telespectators take activaly part to the TV broadcast.
The final objective of this project is to control the animations of an avatar
that should be inserted within a TV show beside the presenter or actor.
By now, the major scenario that we rely on, corresponds to some educational game where a pupil
controls her or his avatar which interacts in a broadcast program led by a ``real'' TV animator.
During such kind of scenarios, commands interaction like pointing, selecting and moving, as well as natural gesture animation, are required.

The paper is organized as following.
Section 2 presents the study held on gestures
semantics, sources, and taxonomies and include the presentation of the free
icons library created and used.
Section 3 shows the technical details of handling the camera input for gestures recognition.
Section 4 shows the modes where the gestural informations generated are used
with some screenshots of the running application.
Section 5 answers the usability and naturelness question of the use of gestures
specially in an ubiquitous environment.
Finally section 6 discusses the future of the work and the possible integration of
other multimodal modules.

\section{Gesture Semantics}
Symbolic gestures, such as emblems, play an important role in human
communication as they can fully take the place of words. These gestures are processed by the same area of the human brain as the spoken language \cite{SymbolicGest}. Hence, these gesures do not come alone, and they are often combined with speech, or with cognitive
activities, as they are classically used to add more precision on the details of a talk.

@@

We have extracted some semantics of such type of gestures, by the analysis of the movements of
a user explaining a story, and by searching for a relation between gestures and the
part of the story being told \cite{gestureThought}. After extraction, most of gestures semantics can be divided into two parts for
their possible future use as:
\begin{description}
 \item[Animation Gestures] used to animate a virtual avatar.
 \item[Command Gestures] used to launch a specific predefined action.
\end{description}

@@

\subsection{Gestures taxonomies}
A rapid look the possible gestures human can perform \cite{Gesturecraft},
lead to the conclusion that the amount of possibilities is tremendous.
Nonetheless, the gestures can be classified according to various criteria not
only to simplify their recognition, but also to allow their reproducing on an
avatar to putting tags on them for future analysis\cite{gestureRecreation}.
Here, we consider a taxonomy based on the context of the gestures as well as on their types:
%%%%%%%
%By looking at the possible gestures human can generate \cite{Gesturecraft}, we
%get a huge amount of possibilities which depend on the context and on the type.
%So, there is a need for a taxonomy.
%Many possible taxonomies exist according to the technical or analytical aspects:
%In order to describe and analyse each gesture, we can rely on the following aspects:
\ignore{\begin{description}
 \item[Gesture Creation Space:] We use this criteria to distinguish between gestures made in the 2D or 3D space.
 \item[Number of used strokes:] We use this criteria to distinguish between gestures constructed by only one path and between those using many paths, so the recognition needs to wait for all strokes before the recognition
 \item[Gesture context:] We use this criteria to distinguish gesture based on the specific use of it, gestures for disabled people, gestures in talk-shows, gestures in meetings.
\end{description}}
%Gesture and Thoughts, page 18
According to McNail \cite{gestureThought}, it is possible and useful to
distinguish between these gestures. Here follows a possible taxonomy :
\begin{description}
 \item[Gesticulation] is a motion that embodies a meaning reliable to the
accompanying speech. It is made chiefly with the arms and hands but is not
restricted to these body parts.
 \item[Speech-linked gestures] are parts of sentences themselves, the gesture
completes the sentence structure.
 \item[Emblems] are conventionalized signs, such as thumbs-up or the ring (first
finger and thumb tips touching, other fingers extended) for ``OK.''
 \item[Pantomime] is dumb show, a gesture or sequence of gestures conveying a
narrative line, with a story to tell, produced without speech.
\end{description}

In our work, we focus on Emblems for Command Gestures, and on Gesticulations for Animation Gestures.
%We have chosen to analyze gesticulations as they are the most frequent.
%, and project them on the technical taxonomies.
%And we have also made our own extraction from Talk-Shows.
%Nous avons choisi celle-ci, car elle correspond aux besoins que nous avons
%identifié dans les scenarios d'utilisation de la télé interactive

@@

% Il faut rédiger complétement cette partie...
%C'est pas compréhensible, et il n'y a pas de verbe dans la seconde phrase.
In order to identify relevant gestures we have analyzed some talk-shows, and we have extracted the most frequent gestures being performed.
According to our scenario of use in the context of interactive TV (RevTV project), we need some
gesture to trigger commands, aas well as other gestures to implement natural interaction with the avatar.
We have created a vector based gestures icons to represent them.
These 4 icons, represented in figure \ref{handg}, are part of the \textbf{ishara} library \cite{ishara}.  They
represent the gestures supported by the software of our application. We provide a movement mode, a multitouch mode, a command mode, a hand fingers recognition,
and, finally, a cursor movement mode (see Section 5 for detailled explanation of these modes).
%and defining a benchmark scenario.
%We have tried to target the following kinds of gestures in a first step.

\png{handg}{Family of supported gestures represented using the free gestures icons library from \cite{ishara}}

%\pagebreak
These five basic gestures modes are already supported in our system as a first step. They will be extended depending on the context fulfilling the requirements
of our scenario.

@@

Figure \ref{pipeline} is a description of the pipeline of the modules used to
extract gestural informations. Each part is described in details in the next
subsections.

@@

\subsection{Camera Input}
We have chosen a low-cost camera that can be integrated with other components in the future.
The PS3Eye camera matched our expectations with a 640x480 resolution combined with a 60 frames-per-second capability.
This camera costs about 40\$, which is a reasonable price for a potential market deployment.

\subsection{Colors Transformations} %%%%%%%%%%%%%%%%%%%
The video processing which depend on the OpenCV library \cite{openCVlib} is based on several modules, such as illustrated in figure \ref{pipeline}.
First, a basic but adaptive skin color segmentation of a large interval of human skin varieties is performed \cite{skinColorSeg}.
We also perform histogram equalization technique to better enlarge the colors to all the space available.
Then we transform the color space from BGR, which come from the camera to YCrCb,
as it is the most adapted one to skin segmentation. Other color spaces like HSV or CIE-Lab can be used for skin color segmentation,
but the transformation from BGR either takes more time (the transformation we use is linear) or needs more parameters
in the next step \cite{skinColorSeg}.

@@

\subsection{Blobs Tracking}
After having the hands and head blobs identified, an efficient
tracking method based on appearance model is used \cite{app06}. Let us note that it handles occlusion problems.
We have used the cvblob implementation \cite{cvblob} for that algorithm which
stores the measures of bounding boxes distance and provides tracks instead of just blobs. %distantBlobTrack.

Once the matrix of distance is defined, the following loops
are executed to handle blobs changes:
\begin{itemize}
 \item A loop to detect inactive tracks: those tracks with no blobs near.
 \item A loop to detect and create new tracks: those blobs without a track near.
 \item A loop to assign blobs to tracks. It makes clusters with blobs that are close and assign them to a track.
In this step some tracks could merge in one.
 \item A last loop which check all inactive tracks to delete the old ones.
\end{itemize}

To better handle blobs, the bounding path of the blobs
is simplified to a set of few points when needed.
This is used to identify picks for the counter mode (see Section 5). Finally, a phase of calibration is triggered when the logic of tracking is lost or damaged.

@@

% inutile : on répète ce qui a déjà été dit !
%We have used four modes in a first core with the possibility to add more sub-gestures in the future.
%This core will be used to control a children game having to select, choose, and answer basic questions.
%We have stick to that scenario to produce these gestures.

@@

\begin{figure}[!htb]
\centering
  \includegraphics[width=0.45\textwidth]{./pics/point.png}
  \includegraphics[width=0.4\textwidth]{./pics/simplify.png}
  \includegraphics[width=0.46\textwidth]{./pics/fingers.png}
  \includegraphics[width=0.46\textwidth]{./pics/move.png}
  \caption{A sub-list of recognition modes showing pointing, number of hand
fingers counting and the movement mode}
 \label{fig:Gestes}
\end{figure}

@@

Spatial gestural interactions always lack from the feedback for the interaction.
This affects also the usability because the user can no more be guided in space
as he was with surface movement or simple physical joysticks.

\section{System integration of this interaction and future work}
Human brain Broca's area, identified as the core of the brain's language
system, are not in fact committed to language processing, but may function as a
modality-independent semiotic system that plays a broader role in human
communication, linking meaning with symbols whether these are words, gestures,
images, sounds, or objects \cite{SymbolicGest}.
This natural handling of multimodal communications gives an argument to build
recognizer taking part of most of these modalities specially to remove ambiguous
gesture meaning decision cases.

@@

\section*{Conclusion}
In this paper we have shown a proof-of-concept mechanism to recognize hand
gestures then use them to control a TV or to take actions in a Game scenario
for pupils.
We have also provided a free library which can be used to represent gestures
in other scenarios and other studies.

The evolution of our system could be possible in the direction of a multimodal
environment in the case where we can be aware of the limits and myths
\cite{Oviatt99tenmyths}.
But this evolution is a natural choice as new devices are getting more
computational power and the ubiquitous environment is becoming a reality.
